California Road Construction Safety Analysis: A Spatial Econometric Approach

### 2025-03-26 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Acknowledged data duplication

# Join accidents to construction buffers - this operation can be computationally intensive
# For very large datasets, consider using a spatial database or partitioning the data
accidents_in_buffers <- st_join(
  df.acc.sf %>% 
    select(accident_id = ID, accident_date = Start_Time, Severity),
  const_buffers %>% 
    select(buffer_id, construction_id = ID, start_date, end_date, is_urban),
  join = st_intersects
)

@@

# Display the first few rows of the joined data
head(accidents_in_buffers %>% st_drop_geometry())
```

>>>>

# Join accidents to construction buffers - this operation can be computationally intensive
accidents_in_buffers <- st_join(
  df.acc.sf %>% 
    select(accident_id = ID, accident_date = Start_Time, Severity),
  const_buffers %>% 
    select(buffer_id, construction_id = ID, start_date, end_date, is_urban),
  join = st_intersects
)

@@

# Display the first few rows of the joined data
head(accidents_in_buffers %>% st_drop_geometry())

# Print number of rows
print(nrow(accidents_in_buffers))
```

As it is visible now, the data almost quadruples for accidents in buffers. Preserving all spatial intersections between accidents and construction buffers through an unrestricted join (`st_intersects`) provides a more realistic representation of road safety dynamics, particularly in complex environments where construction activities frequently overlap. By maintaining the many-to-many relationship, our analysis acknowledges that a single accident may be legitimately associated with multiple construction zones, especially in urban corridors where projects occur in close proximity. This approach aligns with our research objective of measuring the overall safety impact of construction presence rather than attributing causality to specific projects. The difference-in-differences framework naturally accommodates this data structure by focusing on the binary treatment state (construction present or absent) within each spatial-temporal unit, while our fixed effects for buffer zones and time periods effectively control for systematic variations in accident frequencies across locations and periods. This methodological approach captures the real-world density patterns of construction activities without requiring artificial deduplication that might obscure important spatial relationships in our road safety analysis.
